No general trend for multiple dimensions of functional diversity in bird and mammal communities

### 2024-05-30 analysis/06_study_change.Rmd
Minor changes after going over reviewer comments again.

```{r}
change <-  change_groups %>% 
  mutate(study_id = as.integer(study_id)) %>%
  filter(if_any(all_of(c("S", "Jaccard_base", "SES_FRic", "SES_FEve", "SES_FDiv")), ~. != 0))

no_change <- study_meta %>% filter(!study_id %in% change$study_id, study_id %in% model_output$study_id)

redundancy <- change_groups %>% 
  mutate(study_id = as.integer(study_id)) %>%
  filter(S == "neg", SES_FRic == "0", SES_FDiv == "0", SES_FEve == "0") %>%
  left_join(study_meta)

redundancy_gain <- change_groups %>% 
  mutate(study_id = as.integer(study_id)) %>%
  filter(S == "pos", SES_FRic == "0", SES_FDiv == "0", SES_FEve == "0") %>%
  left_join(study_meta)

evenness <- change_groups %>% 
  mutate(study_id = as.integer(study_id)) %>%
  filter(across(!study_id & !SES_FEve, ~. == 0)) %>%
  left_join(study_meta)

increase <- change_groups %>% 
  mutate(study_id = as.integer(study_id)) %>%
  filter(Jaccard_base == "neg", S == "pos", SES_FRic == "pos", SES_FDiv == "0", SES_FEve == "0") %>%
  left_join(study_meta)
```

>>>>

```{r}
# change <-  change_groups %>% 
#   mutate(study_id = as.integer(study_id)) %>%
#   filter(if_any(all_of(c("S", "Jaccard_base", "SES_FRic", "SES_FEve", "SES_FDiv")), ~. != 0))
# 
# no_change <- study_meta %>% filter(!study_id %in% change$study_id, study_id %in% model_output$study_id)
# 
# redundancy <- change_groups %>% 
#   mutate(study_id = as.integer(study_id)) %>%
#   filter(S == "neg", SES_FRic == "0", SES_FDiv == "0", SES_FEve == "0") %>%
#   left_join(study_meta)
# 
# redundancy_gain <- change_groups %>% 
#   mutate(study_id = as.integer(study_id)) %>%
#   filter(S == "pos", SES_FRic == "0", SES_FDiv == "0", SES_FEve == "0") %>%
#   left_join(study_meta)
# 
#   
# evenness <- change_groups %>% 
#   mutate(study_id = as.integer(study_id)) %>%
#   filter(across(!study_id & !SES_FEve, ~. == 0)) %>%
#   left_join(study_meta)
# 
# increase <- change_groups %>% 
#   mutate(study_id = as.integer(study_id)) %>%
#   filter(Jaccard_base == "neg", S == "pos", SES_FRic == "pos", SES_FDiv == "0", SES_FEve == "0") %>%
#   left_join(study_meta)
```

### 2024-05-30 paper/GEB Submission/manuscript_GEB.Rmd
Minor changes after going over reviewer comments again.

Functional diversity offers a potentially powerful addition to species-based approaches for detecting and describing community change by capturing changes in the traits that underpin fundamental ecological processes [@lavorel2002; @mcgill2006; @suding2008].
Functional diversity metrics capture complimentary aspects of the trait space including the volume of the multidimensional space (functional diversity), how species and abundance are distributed across the space (functional evenness and divergence), and where the space occurs across the range of possible trait distributions (community weighted trait means).
Considered together, they describe the impacts of changes in species identity on the building blocks of stability, function, and resilience.

@@

`r colorize("The expectation for functional diversity change across communities is not obvious from past work, and may or may not follow species richness trends", "blue")` [@dornelas2023; @larson2021].
`r colorize("While loss of functional diversity is frequently cited as one of the most pressing concerns of the anthropocene ", "blue")`[@cardinale2012; @dirzo2014; @young2016], observational trends show significant functional loss in some studies [@flynn2009] and no loss even in the most heavily impacted communities for others [@edwards2013; @larsen2018; @matuoka2020].
`r colorize("Broad-scale studies are limited, but show for example functional richness increases for North American birds", "blue")` [@barnagaud2017; @jarzyna2016].
`r colorize("We have stronger expectations for changes in the prevalence of some individual traits. For example, animal body size is expected to decrease as a result of climate change, a phenomena that has been documented in multiple taxa empirically and experimentally", "blue")` [@caruso2014; @forster2012; @huss2019; @sheridan2011; @tseng2018].
`r colorize("For dietary traits, recent work documenting insect declines", "blue")` [@wagner2021] `r colorize("points to potentially significant negative impacts on insectivorous animals", "blue")` [@canaday1996; @sekercioglu2002; @tallamy2021].
`r colorize("Predicted extinctions based on species-level vulnerability point to further dietary shifts, favoring increases in invertivorous species ", "blue")`[@cooke2019].
`r colorize("Some systems also show significant shifts in the prevalence of different kinds of foragers in birds, for example loss of arboreal foragers in agricultural systems", "blue")`[@bain2020]`r colorize(", and loss of neotropical understory foragers even in protected areas", "blue")` [@pollock2022].

@@

`r colorize("The next largest group of studies (7 studies) were those with no species richness trend that showed a trend in at least one functional diversity metric. Based on a species approach alone, these communities apear to either be experiencing no change or some change in species identity, with unknown implications for conservation concern or ecological process. However, with the addition of a functional perspective we found multiple kinds of underlying functional change including increases and decreases in corrected functional richness and divergence and loss of corrected functional evenness. We therefore found evidence of both increases and decreases in the functional capacity of communities independent of species-based metrics, reinforcing the importance of moving beyond soley species-based approaches for biodiversity assessment", "blue")` [@hillebrand2017; @santini2017].

`r colorize("9 studies showed loss of functional redundancy. These studies are characterized by a loss of species richness with no shift in functional diversity metrics, indicating the lost species were functionally similar enough to another species to not impact the overall functional structure. Redundancy is critical for maintaining ecological processes in the face of disturbance as it ensures  functional roles are maintained", "blue")` [@biggs2020]`r colorize(", and a loss of functional redundancy likeley reduces a community's capacity to respond to future stressors ", "blue")`[@reich2012]`r colorize(". These communities are actually fairing better than expected looking at species-based metrics alone, but are also in a precarious position for maintaining ecological function into the future", "blue")` [@pimiento2020a].

>>>>

Functional diversity offers a potentially powerful addition to species-based approaches for detecting and describing community change by capturing changes in the traits that underpin fundamental ecological processes [@lavorel2002; @mcgill2006; @suding2008].
Functional diversity metrics capture complimentary aspects of the trait space including the volume of the multidimensional space (functional diversity), how species and abundance are distributed across the space (functional evenness and divergence), and where the space occurs across the range of possible trait distributions (community weighted trait means).
Considered together, they describe the aspects of the trait space that mediate ecological process.

@@

`r colorize("The expectation for functional diversity change across communities is not obvious from past work, and may or may not follow species richness trends", "blue")` [@dornelas2023; @larson2021].
`r colorize("While loss of functional diversity is frequently cited as one of the most pressing concerns of the anthropocene ", "blue")`[@cardinale2012; @dirzo2014; @young2016], observational trends show significant functional loss in some studies [@flynn2009] and no loss even in the most heavily impacted communities for others [@edwards2013; @larsen2018; @matuoka2020].
`r colorize("Broad-scale observational studies are limited, but show for example functional richness increases for North American birds", "blue")` [@barnagaud2017; @jarzyna2016].
`r colorize("We have stronger expectations for changes in the prevalence of some individual traits. For example, animal body size is expected to decrease as a result of climate change, a phenomena that has been documented in multiple taxa empirically and experimentally", "blue")` [@caruso2014; @forster2012; @huss2019; @sheridan2011; @tseng2018].
`r colorize("For dietary traits, recent work documenting insect declines", "blue")` [@wagner2021] `r colorize("points to potentially significant negative impacts on insectivorous animals", "blue")` [@canaday1996; @sekercioglu2002; @tallamy2021].
`r colorize("Predicted extinctions based on species-level vulnerability point to further dietary shifts, favoring increases in invertivorous species ", "blue")`[@cooke2019].
`r colorize("Some systems also show significant shifts in the prevalence of different kinds of foragers in birds, for example loss of arboreal foragers in agricultural systems", "blue")`[@bain2020]`r colorize(", and loss of neotropical understory foragers even in protected areas", "blue")` [@pollock2022].

@@

`r colorize("The next largest group of studies (7 studies) were those with no species richness trend that showed a trend in at least one functional diversity metric. Based on a species approach alone, these communities apear to either be experiencing no change or some change in species identity, with unknown implications for conservation concern or ecological process. However, with the addition of a functional perspective we found multiple kinds of underlying functional change including increases and decreases in corrected functional richness and divergence and loss of corrected functional evenness. We therefore found evidence of both increases and decreases in the functional capacity of communities independent of species-based metrics, reinforcing the importance of moving beyond soley species-based approaches for biodiversity assessment", "blue")` [@hillebrand2017; @santini2017]. All but one of the studies in this group experienced significant turnover, which was also reflected in trends for multiple *CWM*'s, indicating that the overall functional space was shifting even when some characteristics of the functional space were maintained.

`r colorize("9 studies showed loss of functional redundancy. These studies are characterized by a loss of species richness with no shift in functional diversity metrics, indicating the lost species were functionally similar enough to another species to not impact the overall functional diversity. Redundancy is critical for maintaining ecological processes in the face of disturbance as it ensures  functional roles are maintained", "blue")` [@biggs2020]`r colorize(", and a loss of functional redundancy likeley reduces a community's capacity to respond to future stressors ", "blue")`[@reich2012]`r colorize(". These communities are actually fairing better than expected looking at species-based metrics alone, but are also in a precarious position for maintaining ecological function into the future", "blue")` [@pimiento2020a].
